Resolving gut microbiome networks within Chiropterans

### 2025-07-07 phasegenome_out/doc/paper/draft_manuscript.rmd
update

### Statistical analysis
  Results were imported into R for analysis and visualization using the R packages data.table (**CITATION**), tidyverse (**CITATION**), ggplot (**CITATION**), vegan (**CITATION**), hillR (**CITATION**), phyloseq (**CITATION**), ggtext (**CITATION**), and ggpubr (**CITATION**). mOTUs identified by SingleM were used to assess alpha and beta diversity. Differences in prokaryotic community composition among bat species were analyzed using distance-based redundancy analysis (db-RDA) on a Hellinger distance matrix. Statistical significance was evaluated with 9,999 permutations using the R package vegan (**V**) (**Oksanen et al. 2015**). Richness (Chao1), evenness (Pielou's index), and alpha diversity (Hill numbers at q = 1, corresponding to the inverse Shannon index) were calculated. Comparisons of alpha diversity across bat species were performed using two-way ANOVA.
  To identify microbial mOTUs associated with each bat species, we performed an indicator species analysis following the approach of **De Cáceres and Legendre (2009)** and **De Cáceres et al. (2010)**. Group-equalized correlation coefficients (r₉) were calculated for each bat species to account for differences in taxon abundance and enable direct comparisons among taxa with varying relative abundances. Statistical significance was assessed through 9,999 permutations. All analyses were carried out using the R package indicspecies (**V**) (**De Cáceres et al. 2016**). Subsequently, the relative abundances of indicator mOTUs were calculated for each bat species.

@@

# After rarifaction
#Calling in phyloseq object
rel_abun_phy<-readRDS("../../data/results/tables/singleM__rarefied_phy.rds")

@@

```
  A total of **17702** potential viral sequences were ID from the 6 individual assemblies. Clustering at 95% ANI over 85% of the shortest sequence identified 16289 viral operational taxanomic units (vOTUs) that were $\geq$ 1 kbp in length. Further filtering for sequences $\geq$ 2.5 kpb resulted in a final set of 6235 vOTUs. After rarefaction, `r ncol(vOTU_rar_cov[,-1])` vOTUs were retained. CheckV was used to assess the quality of these sequences, revealing that `r nrow(vOTU_quality[vOTU_quality$completeness>=45.5])` (`r round(nrow(vOTU_quality[vOTU_quality$completeness>=45.5])/nrow(vOTU_quality)*100)`%) were $\geq$ 50% complete including `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Complete"])` complete vOTUs that identified on the bases of direct terminal repeats (DTR), `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5])` high quality vOTUs that were identified on the bases of AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs), `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5])` medium quality vOTUs that were idenified on the bases on AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI-based",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs), and `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5])` low quality vOTUs that were identified based on AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI-based",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs). The reset of the vOTUs (`r nrow(vOTU_quality[vOTU_quality$completeness<45.5|is.na(vOTU_quality$completeness)])`) were of low quality (`r nrow(vOTU_quality[vOTU_quality$completeness<45.5])`) or the quality was undetermined (`r nrow(vOTU_quality[is.na(vOTU_quality$completeness)])`). An unclassified order of the class _Caudoviricetes_, families _Retroviridae_, _Adintoviridae_, and _Iridoviridae_, an unclassified family of _Kyanoviridae_, families _Inoviridae_ and _Bornaviridae_, an unclassified family of _Herelleviridae_, families _Mimiviridae_ and _Parvoviridae_ were the most predominate viral taxa across the gut virome with all 3 of the bat species (**Figure 4**). No statistically significant differences were found in the viral richness, envenness, $\alpha$-diversity, or $B$-diversity nor did indicator analysis reveal any indicator viral spices for the bat species.

@@

### MAG resistome diversity
  Across the gut MAG associated resistome of the 3 bat species, Macaroline resistance was the most abundant AMR class in both _M. crenulatum_ and _P. discolor_ while Tetracycline were the most abundant in _P. hastatus_. The second most prevalent AMR varied by bat species, with Tertracycline resistance observed in _M. crenulatum_, Glycopeptide resistance observed in _P. discolor_, and Macrolide resistance abserved in _P. hastatus_ (**Fig X**). The most abundant gene associated with Macrolide resistance was abc-f (all three bat species), _mef(B)_ ( _P. hastatus_ ), _erm_ ( _M. crenulatum_ ), and _estT_ ( _M. crenulatum_ and _P. discolor_). Tetracycline resistance was dominated by _tet_ genes ( _tet_, _tetA(46)_, and _tetB(46)_) within all three bat species while _mepA_ was more abundant in both _P. hastatus_ and _P. discolor_. The most predominate genes associated with Glycopeptide resistance was represented _vanR_, _vanS_, _vanH_, and _vanR-B_ in the guts across all three bat species while _vanY-D_ was highly abundant in the guts of _M. crenulatum_ and _P. discolor_.

  Within the MAGs, a total of **52** AMR genes across **X** AMR classes were unique to _M. crenulatum_, **22** AMR genes across **X** AMR classes were unique to _P. hastatus_, while no AMR genes were found to be unique to _P. discolor_ (**Fig X**). Unique AMR genes within _M. crenulatum_ were within the classes **X,Y,andZ**. The most abundant of these were **A,B,andC**. The **22** AMR genes unique to  _P. hastatus_ were within the classes **X,Y,andZ** with the most predominate genes being **A,B,andC**. Interesting **71** AMR genes (representing **X** classes) were shared between _M. crenulatum_ and _P. hastatus_ while only 3 (representing **X** classes) were shared between _P. discolor_ and _P. hastatus_ and only 2 between (representing **X** classes) _M. crenulatum_ and _P. discolor_.

@@

### vMAG resistome diversity
  Within the gut vMAG associated resistome of the 3 bat species, Macaroline resistance was the most abundant AMR class in both _M. crenulatum_ and _P. discolor_. In fact, this was the only class of antimicrobial resistance found within _P. discolor_. The Mupirocin resistance class was the most abundant in _P. hastatus_. The second most prevalent AMR varied between _M. crenulatum_ and _P. hastatus_, with Quinolone resistance observed in _M. crenulatum_ and  Macaroline resistance observed in _P. hastatus_ (**Fig X**). The only AMR gene associated with Macrolide resistance was _abc-f_ (all three bat species). Quinolone resistance was represented by _qnrVC_, _qnrE_, _qnrS_, _qnrA_, and _qnrB_ within _M. crenulatum_ and by qnrD in _P. hastatus_.

  A total of **11** AMR genes across **X** AMR classes were unique to _M. crenulatum_, **5** AMR genes across **X** AMR classes were unique to _P. hastatus_, while no AMR genes were found to be unique to _P. discolor_ (**Fig X**). Unique AMR genes within _M. crenulatum_ were within the classes **X,Y,andZ**. The most abundant of these were **A,B,andC**. The **5** AMR genes unique to  _P. hastatus_ were within the classes **X,Y,andZ** with the most predominate genes being **A,B,andC**. Interesting **7** AMR genes (representing **X** classes) were shared between _M. crenulatum_ and _P. hastatus_ while none were shared between _P. discolor_ and _P. hastatus_ nor between _M. crenulatum_ and _P. discolor_.

>>>>

### Statistical analysis
  Results were imported into R for analysis and visualization using the R packages data.table (**CITATION**), tidyverse (**CITATION**), ggplot (**CITATION**), vegan (**CITATION**), hillR (**CITATION**), phyloseq (**CITATION**), ggtext (**CITATION**), and ggpubr (**CITATION**). mOTUs identified by SingleM were used to assess $a$- and β-diversity. Differences in prokaryotic community composition among bat species were analyzed using distance-based redundancy analysis (db-RDA) on a Hellinger distance matrix. Statistical significance was evaluated with 9,999 permutations using the R package vegan (**V**) (**Oksanen et al. 2015**). Richness (Chao1), evenness (Pielou's index), and alpha diversity (Hill numbers at q = 1, corresponding to the inverse Shannon index) were calculated. Comparisons of alpha diversity across bat species were performed using two-way ANOVA.
  To identify microbial mOTUs associated with each bat species, we performed an indicator species analysis following the approach of **De Cáceres and Legendre (2009)** and **De Cáceres et al. (2010)**. Group-equalized correlation coefficients (r₉) were calculated for each bat species to account for differences in taxon abundance and enable direct comparisons among taxa with varying relative abundances. Statistical significance was assessed through 9,999 permutations. All analyses were carried out using the R package indicspecies (**V**) (**De Cáceres et al. 2016**). Subsequently, the relative abundances of indicator mOTUs were calculated for each bat species.

@@

# After rarifaction
#Calling in phyloseq object
rel_abun_phy<-readRDS("../../data/results/singleM__rarefied_phy.rds")

@@

```
  A total of **17702** potential viral sequences were ID from the 6 individual assemblies. Clustering at 95% ANI over 85% of the shortest sequence identified 16289 viral operational taxanomic units (vOTUs) that were $\geq$ 1 kbp in length. Further filtering for sequences $\geq$ 2.5 kpb resulted in a final set of 6235 vOTUs. After rarefaction, `r ncol(vOTU_rar_cov[,-1])` vOTUs were retained. CheckV was used to assess the quality of these sequences, revealing that `r nrow(vOTU_quality[vOTU_quality$completeness>=45.5])` (`r round(nrow(vOTU_quality[vOTU_quality$completeness>=45.5])/nrow(vOTU_quality)*100)`%) were $\geq$ 50% complete including `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Complete"])` complete vOTUs that identified on the bases of direct terminal repeats (DTR), `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5])` high quality vOTUs that were identified on the bases of AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs), `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5])` medium quality vOTUs that were idenified on the bases on AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI-based",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs), and `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5])` low quality vOTUs that were identified based on AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI-based",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs). The reset of the vOTUs (`r nrow(vOTU_quality[vOTU_quality$completeness<45.5|is.na(vOTU_quality$completeness)])`) were of low quality (`r nrow(vOTU_quality[vOTU_quality$completeness<45.5])`) or the quality was undetermined (`r nrow(vOTU_quality[is.na(vOTU_quality$completeness)])`). An unclassified order of the class _Caudoviricetes_, families _Retroviridae_, _Adintoviridae_, and _Iridoviridae_, an unclassified family of _Kyanoviridae_, families _Inoviridae_ and _Bornaviridae_, an unclassified family of _Herelleviridae_, families _Mimiviridae_ and _Parvoviridae_ were the most predominate viral taxa across the gut virome with all 3 of the bat species (**Figure 4**). No statistically significant differences were found in the viral richness, envenness, $\alpha$-diversity, or β-diversity nor did indicator analysis reveal any indicator viral spices for the bat species.

@@

### MAG resistome diversity
  Across the gut MAG associated resistome of the 3 bat species, Macaroline resistance was the most abundant AMR class in both _M. crenulatum_ and _P. discolor_ while Tetracycline were the most abundant in _P. hastatus_. The second most prevalent AMR varied by bat species, with Tertracycline resistance observed in _M. crenulatum_, Glycopeptide resistance observed in _P. discolor_, and Macrolide resistance abserved in _P. hastatus_ (**Fig 7**). The most abundant gene associated with Macrolide resistance was abc-f (all three bat species), _mef(B)_ ( _P. hastatus_ ), _erm_ ( _M. crenulatum_ ), and _estT_ ( _M. crenulatum_ and _P. discolor_). Tetracycline resistance was dominated by _tet_ genes ( _tet_, _tetA(46)_, and _tetB(46)_) within all three bat species while _mepA_ was more abundant in both _P. hastatus_ and _P. discolor_. The most predominate genes associated with Glycopeptide resistance was represented _vanR_, _vanS_, _vanH_, and _vanR-B_ in the guts across all three bat species while _vanY-D_ was highly abundant in the guts of _M. crenulatum_ and _P. discolor_.

  Within the MAGs, a total of **52** AMR genes across **15** AMR classes were unique to _M. crenulatum_, and **22** AMR genes across **8** AMR classes were unique to _P. hastatus_, while no AMR genes were found to be unique to _P. discolor_ (**Fig. X**). Although these AMR genes were unique to either _M. crenulatum_ or _P. hastatus_, **7** of the **8** AMR gene classes present in _P. hastatus_ (Aminoglycoside, β-lactam, Trimethoprim, Macrolide, Fosfomycin, Lincosamide, and Colistin resistance) were also found in _M. crenulatum_. Only the Streptothricin class was unique to _P. hastatus_. AMR classes unique to _M. crenulatum_ included Bleomycin, Phenicol, Lincosamide/Macrolide, Phenicol/Oxazolidinone, Sulfonamide, Tetracycline, Glycopeptide, and Streptogramin resistance.

  Among the unique AMR genes, the most prominent in _M. crenulatum_ were _acc(2')-lc_, _aac(6')-Il_ (Aminoglycoside), _dfrF_ (Trimethoprim), _blaDHA_, _blaGMB_, _blaMAB_, _blaBUT_, _blaOCH_, _cfiA_ (β-lactam), and _vanZ-Pt_ (Glycopeptide). In contrast, the most predominant AMR genes in _P. hastatus_ were _aac(3)-VIII_, _aac(6')-I_, _aac(6')-Ian_, _aacA10_, _aph(3')-V_, _aph(7')-Ia_, _aph(9)-Ia_ (Aminoglycoside), _dfrA36_, _dfrG_ (Trimethoprim), _blaPDC_, _blaZ_ (β-lactam), _mcr-10_ (Colistin), _fosC_ (Fosfomycin), _erm(55)_, _mph(C)_ (Macrolide), and _sat2_ (Streptothricin). All other AMR genes had abundances of **< 5** TPM.

@@

### vMAG resistome diversity
  Within the gut vMAG associated resistome of the 3 bat species, Macaroline resistance was the most abundant AMR class in both _M. crenulatum_ and _P. discolor_. In fact, this was the only class of antimicrobial resistance found within _P. discolor_. The Mupirocin resistance class was the most abundant in _P. hastatus_. The second most prevalent AMR varied between _M. crenulatum_ and _P. hastatus_, with Quinolone resistance observed in _M. crenulatum_ and  Macaroline resistance observed in _P. hastatus_ (**Fig 8**). The only AMR gene associated with Macrolide resistance was _abc-f_ (all three bat species). Quinolone resistance was represented by _qnrVC_, _qnrE_, _qnrS_, _qnrA_, and _qnrB_ within _M. crenulatum_ and by qnrD in _P. hastatus_.

  A total of **11** AMR genes across **6** AMR classes were unique to _M. crenulatum_, **5** AMR genes across **5** AMR classes were unique to _P. hastatus_, while no AMR genes were found to be unique to _P. discolor_ (**Fig X**). While these AMR genes are unique to either _M. crenulatum_ or _P. hastatus_, Quinolone, Trimethoprim, and Glycopeptide resistant classes are present in both bat species. AMR classes Phenicol, Rifamycin, and Streptogramin are unique classes within _M. crenulatum_ while Mupirocin and Thiostrepton are unique AMR classes within _P. hastatus_.

  Among the unique AMR genes, the most prominent in _M. crenulatum_ were _qnrA_, _qnrB_, _qnrVC_, _qnrE_, _qnrS_ (Quinolone), _dfrA_ (Trimethoprim), and _vat_ (Streptogramin). In contrast, the most predominant AMR genes in _P. hastatus_ were _mupA_ (Mupirocin), _dfra35_ (Trimthoprim), _nshR_ (Thiostrepton), _vanY_ (Glysopeptide), and _qnrD_ (Quinolone). All other AMR genes had abundances of **< 5** TPM.
